Data Science Applied to Ag - Final Project - ML

### 2025-04-26 code/Draft_script.qmd
d

fe_month_sum <- fe_month %>%
  group_by(year, site, hybrid, lon, lat, month_abb, yield, prev_crop, planting_stages, harvest_stages) %>% #If we do a summarise() after group_by(), any column that's not in the group_by() is gone, so we need to include "strength_gtex" in the group_by() to include it in the data frame because "strength_gtex" is our response variable so we must keep it in the data frame 
  #Because we are gonna be applying a "summarize()" function to different columns, we are gonna use a function called "across()" 
  summarise(across(.cols = c(soil.ph, 
                             soil.om.pct, 
                             soil.k.ppm, 
                             soil.p.ppm,
                             dayl.s,
                             srad.wm2,
                             tmax.c,
                             tmin.c,
                             vp.pa),
                   .fns = mean, #do not indicate the actual function "mean()", just use the word "mean"
                   .names = "mean_{.col}"), #specifying the weather variables that we want their mean as new column variables #1st across() is applying the "mean" function (to summarize "mean")
            across(.cols = prcp.mm,
                   .fns = sum,
                   .names = "sum_{.col}"
                   ) #specifying the weather variable (prcp.mm) that we want its sum as new column variable #2nd across() is summarizing sum #2nd across() is applying the "sum" function (to summarize "sum")
            ) %>%
  ungroup() #To convert from "group" to "tibble"

@@

```{r}
final_spec %>%
  fit(strength_gtex ~ .,
      data = bake(weather_prep, 
                  weather_train)) %>%
  augment(new_data = bake(weather_prep, 
                          weather_train)) %>% 
  rmse(strength_gtex, .pred) %>%
  bind_rows(

# R2
final_spec %>%
  fit(strength_gtex ~ .,
      data = bake(weather_prep, 
                  weather_train)) %>%
  augment(new_data = bake(weather_prep, 
                          weather_train)) %>% 
  rsq(strength_gtex, .pred))
```

@@

```{r}
final_fit %>%
  collect_predictions() %>%
  ggplot(aes(x = strength_gtex,
             y = .pred)) +
  geom_point() +
  geom_abline() +
  geom_smooth(method = "lm") +
  scale_x_continuous(limits = c(20, 40)) +
  scale_y_continuous(limits = c(20, 40)) 
```

@@

```{r final_spec}
final_spec %>%
  fit(strength_gtex ~ .,
         data = bake(weather_prep, weather_train)) %>% #There little change in variable improtance if you use full dataset
    vi() %>%
  mutate(
    Variable = fct_reorder(Variable, 
                           Importance)
  ) %>%
  ggplot(aes(x = Importance, 
             y = Variable)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)
```

@@

Metrics on **train set** (for curiosity and compare to test set):  
```{r}
# RMSE
final_spec %>%
  fit(strength_gtex ~ .,
      data = bake(weather_prep, 
                  weather_train)) %>%
  augment(new_data = bake(weather_prep, 
                          weather_train)) %>% 
  rmse(strength_gtex, .pred) %>%
  bind_rows(

    # R2
    final_spec %>%
      fit(strength_gtex ~ .,
          data = bake(weather_prep, 
                      weather_train)) %>%
      augment(new_data = bake(weather_prep, 
                              weather_train)) %>% 
      rsq(strength_gtex, .pred)

@@

Predicted vs. observed plot:  
```{r}
final_fit %>%
  collect_predictions() %>%
  ggplot(aes(x = strength_gtex,
             y = .pred)) +
  geom_point() +
  geom_abline() +
  geom_smooth(method = "lm") +
  scale_x_continuous(limits = c(20, 40)) +
  scale_y_continuous(limits = c(20, 40)) 
```

@@

final_spec %>%
  fit(strength_gtex ~ .,
         data = bake(weather_prep, weather)) %>%
    vi() %>%
  mutate(
    Variable = fct_reorder(Variable, 
                           Importance)
  ) %>%
  ggplot(aes(x = Importance, 
             y = Variable)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)

>>>>

fe_month_sum <- fe_month %>%
  group_by(year, site, hybrid, lon, lat, month_abb, yield, prev_crop, planting_stages, harvest_stages) %>% #If we do a summarise() after group_by(), any column that's not in the group_by() is gone, so we need to include "yield" in the group_by() to include it in the data frame because "yield" is our response variable so we must keep it in the data frame 
  #Because we are gonna be applying a "summarize()" function to different columns, we are gonna use a function called "across()" 
  summarise(across(.cols = c(soil.ph, 
                             soil.om.pct, 
                             soil.k.ppm, 
                             soil.p.ppm,
                             dayl.s,
                             srad.wm2,
                             tmax.c,
                             tmin.c,
                             vp.pa),
                   .fns = mean, #do not indicate the actual function "mean()", just use the word "mean"
                   .names = "mean_{.col}"), #specifying the weather variables that we want their mean as new column variables #1st across() is applying the "mean" function (to summarize "mean")
            across(.cols = prcp.mm,
                   .fns = sum,
                   .names = "sum_{.col}"
                   ) #specifying the weather variable (prcp.mm) that we want its sum as new column variable #2nd across() is summarizing sum #2nd across() is applying the "sum" function (to summarize "sum")
            ) %>%
  ungroup() #To convert from "group" to "tibble"

@@

```{r}
final_spec %>%
  fit(yield ~ .,
      data = bake(weather_prep, 
                  weather_train)) %>%
  augment(new_data = bake(weather_prep, 
                          weather_train)) %>% 
  rmse(yield, .pred) %>%
  bind_rows(

# R2
final_spec %>%
  fit(yield ~ .,
      data = bake(weather_prep, 
                  weather_train)) %>%
  augment(new_data = bake(weather_prep, 
                          weather_train)) %>% 
  rsq(yield, .pred))
```

@@

```{r}
final_fit %>%
  collect_predictions() %>%
  ggplot(aes(x = yield,
             y = .pred)) +
  geom_point() +
  geom_abline() +
  geom_smooth(method = "lm") +
  scale_x_continuous(limits = c(20, 40)) +
  scale_y_continuous(limits = c(20, 40)) 
```

@@

```{r final_spec}
final_spec %>%
  fit(yield ~ .,
         data = bake(weather_prep, weather_train)) %>% #There little change in variable improtance if you use full dataset
    vi() %>%
  mutate(
    Variable = fct_reorder(Variable, 
                           Importance)
  ) %>%
  ggplot(aes(x = Importance, 
             y = Variable)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)
```

@@

Metrics on **train set** (for curiosity and compare to test set):  
```{r}
# RMSE
final_spec %>%
  fit(yield ~ .,
      data = bake(weather_prep, 
                  weather_train)) %>%
  augment(new_data = bake(weather_prep, 
                          weather_train)) %>% 
  rmse(yield, .pred) %>%
  bind_rows(

    # R2
    final_spec %>%
      fit(yield ~ .,
          data = bake(weather_prep, 
                      weather_train)) %>%
      augment(new_data = bake(weather_prep, 
                              weather_train)) %>% 
      rsq(yield, .pred)

@@

Predicted vs. observed plot:  
```{r}
final_fit %>%
  collect_predictions() %>%
  ggplot(aes(x = yield,
             y = .pred)) +
  geom_point() +
  geom_abline() +
  geom_smooth(method = "lm") +
  scale_x_continuous(limits = c(20, 40)) +
  scale_y_continuous(limits = c(20, 40)) 
```

@@

final_spec %>%
  fit(yield ~ .,
         data = bake(weather_prep, weather)) %>%
    vi() %>%
  mutate(
    Variable = fct_reorder(Variable, 
                           Importance)
  ) %>%
  ggplot(aes(x = Importance, 
             y = Variable)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)
